Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-03-22 PAPER.tex
edited jackknife text more

\begin{itemize}
%\item Original Data (black): This is identical to the unweighted, revised PAPER-64 power spectrum in Figure \ref{fig:ps1_data} (one baseline type only) with weighting matrix $\textbf{I}$. There are clear detections at low $k$'s.
\item Baselines: We split our dataset into two halves, where each contains half of the total baselines used in the 
analysis. No baselines are repeated between the two datasets.
\item Sidereal Hour: We split our dataset into two halves based on LST, namely the first half (LSTs 0.5-4.6 hours) and second half (LSTs 
4.6-8.6 hours).
\item Day: We split our dataset into even and odd Julian days. We form power spectra for each separately, only cross-multiplying `even' with `even', for example. Although we don't allow this in our standard pipeline, we include these same-dataset cross-multiplications to allow a power spectrum to be estimated from solely the `even' data (similarly for `odd'). If the same sky signal is in both the `even' and `odd' datasets, we expect it to cancel out. %However, allowing these cross-multiplications can in practice introduce a low level noise bias... 
\end{itemize}

\begin{figure*}
	\centering
	\includegraphics[width=0.47\textwidth]{plots/null_bls.png}
	\includegraphics[width=0.47\textwidth]{plots/null_lsts.png}
	\includegraphics[width=0.47\textwidth]{plots/null_eo.png}
	\caption{Power spectrum results (with $2\sigma$ errors) for three null tests, where a jackknife is taken along the baseline axis (top left), LST axis (top right), and even/odd Julian date axis (bottom). The results shown are unweighted (no signal loss), where the power spectrum values plotted are computed from the difference between two power spectra of a jackknife. The gray shaded region in each plot is the estimated $2\sigma$ theoretical noise limit given the parameters of each test. We find that there are no significant systematics that differ between the tests, though there are low level biases at high-$k$ values in both the LST test (likely due to foregrounds that vary in LST) and even/odd test (likely due to \cc{??}). All three tests are dominated by residual foregrounds at $k$-values near the horizon ($k\sim\pm 0.1$ $h$Mpc$^{-1}$), which is not surprising given that foregrounds are brightest in this region post-delay filtering.}
	\label{fig:null}
\end{figure*}

@@

Investigating Figure \ref{fig:null}, we find that the baselines test is fully consistent with noise (all error bars pass through zero). Additionally, the differenced power spectrum values (black points) all lie within the gray-shaded estimated error region for high $k$'s. Therefore, we conclude that particular baselines are not to be blamed for the bias in our dataset.

For the LST test, $90\%$ of all the error bars pass through zero, which is slightly lower than the expected $95\%$. Additionally, not all power spectrum points lie within the gray-shaded region, even at high-$k$. This LST-dependent systematic is likely caused by residual foregrounds that vary in LST. In other words, the two jackknife halves differ by an amount greater than thermal noise and do not completely cancel out the sky signal when differencing.

The even/odd test shows $81\%$ of all error bars being consistent with zero, with several power spectrum points falling outside the estimated error region. \cc{why is that?} %Rather than attribute these biases to systematics that differ \textit{between} the `even' and `odd' datasets, we believe that they are caused by the squaring of thermal noise (which was allowed in the construction of this test). This supports our choice to not allow the cross-multiplication between the same LST-binned datasets in our full pipeline.

Finally, we note that all three null tests exhibit error bars that are strongly $k$-dependent. The errors are largest just outside the horizon ($k\sim\pm 0.1$ $h$Mpc$^{-1}$) which is expected since foregrounds are brightest in this region post-delay filtering. We also note that all three tests result in power spectrum errors that are larger than the estimated thermal noise, even at high-$k$. There are a couple possible explanations for this. Although we expect each cross-multiplication to have independent noise (except in the even/odd case), there is still the possibility of a noise-foreground coupling term that can introduce power. This cross-term would vary with foreground strength, therefore being consistent with the shape of the results. Alternatively, our estimate of the thermal noise may be too low, as all three tests seem to imply a larger, but consistent, estimate.

>>>>

\begin{itemize}
%\item Original Data (black): This is identical to the unweighted, revised PAPER-64 power spectrum in Figure \ref{fig:ps1_data} (one baseline type only) with weighting matrix $\textbf{I}$. There are clear detections at low $k$'s.
\item Baselines: We split our dataset into two halves, where each contains half of the total baselines used in the 
analysis. No baselines are repeated between the two datasets.
\item Sidereal Hour: We split our dataset into two halves based on LST, namely the first half (LSTs 0.5-4.6 hours) and second half (LSTs 
4.6-8.6 hours).
\item Day: We split our dataset into even and odd Julian days. We form power spectra for each separately, allowing the cross-multiplication of `even' with `even', for example, for this null test only. If the same sky signal is in both the `even' and `odd' datasets, we expect it to cancel out. %However, allowing these cross-multiplications can in practice introduce a low level noise bias... 
\end{itemize}

\begin{figure*}
	\centering
	\includegraphics[width=0.47\textwidth]{plots/null_bls.png}
	\includegraphics[width=0.47\textwidth]{plots/null_lsts.png}
	\includegraphics[width=0.47\textwidth]{plots/null_eo.png}
	\caption{Power spectrum results (with $2\sigma$ errors) for three null tests, where a jackknife is taken along the baseline axis (top left), LST axis (top right), and even/odd Julian date axis (bottom). The results shown are unweighted (no signal loss), where the power spectrum values plotted are computed from the difference between two power spectra produced on either side of the jackknife axis. The gray shaded region in each plot is the estimated $2\sigma$ theoretical noise limit given the parameters of each test. We find that there are no significant systematics that differ between the tests, though there are low level biases at high-$k$ values in both the LST test and even/odd test. All three tests, whether biased or not, have an extra noise bias that is strongly $k$-dependent. Residual foregrounds dominate at $k$-values near the horizon ($k\sim\pm 0.1$ $h$Mpc$^{-1}$), which is not surprising given that foregrounds are brightest in this region post-delay filtering.}
	\label{fig:null}
\end{figure*}

@@

In investigating Figure \ref{fig:null}, we focus on two main aspects of each test - whether the error bars are consistent with zero, and whether they are consistent with thermal noise. The former indicates a `passing' null test where there is no bias along the jackknife axis. The latter reveals whether there is an additional noise bias present that lies above the thermal noise.

We find that the baselines test is fully consistent with noise, as all error bars pass through zero. Therefore, we conclude that particular baselines are not to be blamed for the bias in our dataset. For the LST test, $90\%$ of all the error bars pass through zero, which is slightly lower than the expected $95\%$. This indicates a low level bias that is LST-dependent, and likely caused by residual foregrounds that vary in LST. In other words, the two jackknife halves differ by an amount greater than thermal noise and do not completely cancel out the sky signal when differencing. The even/odd test shows $81\%$ of all error bars being consistent with zero, again indicating the presence of a low level bias. The cause of this particular systematic is not obvious, and requires further investigation that we do not carry out in this paper.

Finally, we note that all three null tests exhibit error bars that are strongly $k$-dependent. The errors are largest just outside the horizon ($k\sim\pm 0.1$ $h$Mpc$^{-1}$) which is expected since foregrounds are brightest in this region post-delay filtering. We also note that all three tests, whether biased or unbiased, result in power spectrum errors that are slightly larger than the estimated thermal noise at high-$k$. This indicates an additive noise component that is increasing our errors (while not affecting whether they pass through zero or not).

There are a couple possible explanations for this. Although we expect each cross-multiplication that is used in power spectrum estimation to have independent noise, there is still the possibility of a noise-foreground coupling term that can introduce power. This is because cross-multiplications produce four additive terms - a signal-squared term (where signal includes both foregrounds and EoR), two cross-terms between the signal and noise, and one noise-only term. When differencing two power spectra, we expect the signal-only term to subtract out for a `passing' null test, and we expect the noise-only terms to be consistent with thermal noise. The cross-terms, however, can introduce a noise bias that would vary with foreground strength, therefore being consistent with the shape of our results. That being said, an alternate explanation for the large error bars seen at high $k$'s could be that our estimate of the thermal noise is too low, as all three tests seem to imply a larger, but consistent, estimate.
